Telemetry Module for Solar Vehicle

### 2021-06-17 documentation/FinalPaper/paper.tex

\section{Design}
The overall system consists of three separate modules; the Solar Car Unit which communicates with the Support Vehicle Unit via a wireless RF connection, which in turn interacts with the Support Vehicle Laptop through a serial USB connection. The schematic for the system is outlined in Figure \ref{fig:schematic}. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{documentation/images/schematic.png}
    \caption{Overview of the system's components in block diagram form.}
    \label{fig:schematic}
\end{figure}

\section{Components}
\subsection{CAN bus}
% MCP
\subsection{RF network} % Steffan
The system uses two nRF24L01+ transceivers to provide a bidirectional connection between the Solar Car and Support Vehicle over the 2.4 GHz ISM band. The module features a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data rates of up to 2 Mbit/s.

An advantage of using an off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^{\copyright}$.

The nRF24L01+ module uses the Enhanced Shockburst $\copyright$ packet structure to transmit messages over the 2.4 GHz ISM band. Messages contain a data payload of up to 32 bytes, allowing us comfortably to transmit one CAN telemetry message per packet \cite{shockburst}. Further, the package structure includes a preamble, address, packet and CRC bytes segments as shown in Figure \ref{fig:shockburst}.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.png}
    \caption{The Enhanced Shockburst $\copyright$ packet structure.}
    \label{fig:shockburst}
\end{figure}

\subsection{SD card}
\subsection{Microcontrollers} % Steffan

@@

Several of the key functionalities of the telemetry are implemented in software. This together with the aim of designing a system which is easy to extend in the future makes the choice of software frameworks very important, since they affect reusability and maintainability of the developed code.

\subsection{Build system}
As a build system, platformIO \cite{platformIO} was chosen for the project. It is a version control friendly, cross-platform embedded build tool which includes library management and works by defining \textit{environments} to allow development with different embedded platforms on the same code base. Other project groups working on the solar car in parallel integrated their software into the platformIO ecosystem as well, making extensive code reuse in the future possible.

@@

\subsection{Other Libraries}
% ACAN.h
% RF24.h
%

@@

\subsection{Software Stack}

In this section the main software tools and libraries used throughout the project are introduced.

\subsubsection{PlatformIO}

\subsubsection{ChibiOS}
An open-source RTOS for embedded systems including Teensy. It supports threads, semaphores, data-sharing and timers and more to ease the scheduling of different tasks by the program kernel.

\subsubsection{ACAN}
The ACAN library by Pierre Molinaro is a CAN network driver for the Teensy 3.6 board, as well as older Teensy series 3 boards. The library has been important for the connection between the CAN network and the ports on the Teensy board. We have only used the classic CAN configuration, but in addition to this, the library supports Flex CAN and CANFD. The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s\cite{ACAN}.

\subsubsection{Other Libraries}

\subsection{CAN} % Victor
The CAN network in the solar car enables the micro-controllers at each electrical unit in the car to send and receive data without a host computer. This is possible with the special structure of the CAN message.

\begin{figure}[h]
    %\centering
    \includegraphics[scale=0.35]{documentation/images/detailed-can-data-frame-architecture.jpg}
    \caption{Structure of a CAN frame}
    \label{fig:CANframe}
\end{figure}

The frame uses an identifier to tell which messages goes to which unit. When two messages tries to access the bus, the one with the highest priority, lowest ID, will be granted access to the bus, while the other message has to wait. After the ID comes the RTR bit which determines whether the frame is set to be at sent or received by the node. The control field, sometimes called DLC, tells the amount of bytes used in the following frame field with actual data. The rest of the fields in fig. \ref{fig:CANframe} is not important concerning the nodes ability to communicate across the CAN network.

The CAN bus itself is two twisted wires, CAN High and CAN Low. These control the bits in the CAN message by setting a dominant and recessive voltage, which corresponds to a 0 and 1 respectively. A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.5V and the CAN Low wire is driven at 2.5V.

@@

\section{Discussion}
% variable length messages would reduce overhead -> but most messages contain CAN and are optimized
%

>>>>

\section{Design}
The overall system consists of three separate modules; the Solar Car Unit which communicates with the Support Vehicle Unit via a wireless RF connection, which in turn interacts with the Support Vehicle Laptop through a serial USB connection. The schematic for the system is outlined in Figure \ref{fig:schematic}. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{documentation/images/schematic.png}
    \caption{Overview of the system's components in block diagram form.}
    \label{fig:schematic}
\end{figure}
%

\section{Components}
\subsection{CAN bus}
The CAN network in the solar car enables the micro-controllers at each electrical unit in the car to send and receive data without a host computer. This is possible with the special structure of the CAN message.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{documentation/images/detailed-can-data-frame-architecture.jpg}
    \caption{Structure of a CAN frame}
    \label{fig:CANframe}
\end{figure}

The frame uses an identifier to tell which messages goes to which unit. When two messages try to access the bus, the one with the highest priority, lowest ID, will be granted access to the bus, while the other message has to wait. After the ID comes the RTR bit which determines whether the frame remotely requests or actually sends data to a node. The control field, sometimes called DLC, holds the amount of bytes used in the following frame field containing the actual data. The rest of the fields in fig. \ref{fig:CANframe} serve error checking and coordination of communication on the bus.

The physical CAN bus itself consists of a differential pair of twisted wires, CAN High and CAN Low. These control the bits in the CAN message by setting a dominant and recessive voltage, which corresponds to a 0 and 1 respectively. A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.5V and the CAN Low wire is driven at 2.5V.

The Teensy 3.6 used in the solar car features a CAN controller handling the communication with the CAN bus. In order to interface with the physical CAN bus though, a transceiver is needed. Here the MCP2551 was chosen, which maintains one input and one output serial stream to the CAN controller on the Teensy and drives the CAN bus when a message is transmitted.
% MCP
\subsection{RF network} % Steffan
The system uses two nRF24L01+ transceivers to provide a bidirectional connection between the Solar Car and Support Vehicle over the 2.4 GHz ISM band. The module features a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data transfer rates of up to 2 Mbit/s.

An advantage of using the off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^\copyright$ \cite{shockburst}. Package structures contain a preamble, address, packet, data payload and CRC bytes segments as shown in Figure \ref{fig:shockburst}.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.png}
    \caption{The Enhanced Shockburst$^\copyright$ packet structure.}
    \label{fig:shockburst}
\end{figure}

\subsection{SD card}

We chose a SD card as a permanent local storage solution for the solar cars "black box" due to its ease of use, its interchangeability and since the Teensy 3.6 has an integrated SD card slot on its PCB. Furthermore, this solution makes it easy to increase the storage capacity to more than sufficient sizes with new SD card sizes exceeding 1 TiB.

\subsection{Microcontrollers} % Steffan
Finally, Teensy 3.6 and 4.0 microcontrollers are used as connection hubs in the Solar Car and Support Vehicle Units respectively. The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data. Being based on the Arduino framework, both Teensys offer support for numerous protocols and software libraries to ease the implementation of our design.

@@

Several of the key functionalities of the telemetry module are implemented in software. This together with the aim of designing a system which is easy to extend in the future makes the choice of software frameworks very important, since they affect reusability and maintainability of the developed code.

\subsection{Build system}
As a build system, PlatformIO \cite{platformIO} was chosen for the project. It is a version control friendly, cross-platform embedded build tool which includes library management and works by defining \textit{environments} to allow development with different embedded platforms on the same code base. Other project groups working on the solar car in parallel integrated their software into the platformIO ecosystem as well, making extensive code reuse in the future possible.

@@

\subsection{Other Libraries}

In order to interface with the components presented in the previous section, a set of libraries were used.

The ACAN library by Pierre Molinaro is a CAN network driver for the Teensy 3.6 board, as well as older Teensy series 3 boards. The library has been important for the connection between the CAN network and the ports on the Teensy board. We have only used the classic CAN configuration, but in addition to this, the library supports Flex CAN and CANFD. The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s\cite{ACAN}.
% ACAN.h
% RF24.h
%

@@

\subsection{CAN} % Victor

@@

\href{https://youtu.be/-iwST3REn40}{video link}
\section{Discussion}
% variable length messages would reduce overhead -> but most messages contain CAN and are optimized
% range test
% throughput limitations
